Results_setup

### 2020-04-22 doc/Results_v2.rmd
fixing bugs and playing with triangle plots

load(file = "~./MLH1repo/data/MLH1/MLH1_data_setup_4.21.20.RData")
#load(file = "C:/Users/alpeterson7/Documents/MLH1repo/data/MLH1/MLH1_data_setup_1.31.20.RData")
#2.2.20
#load(file="C:/Users/April/Documents/MLH1repo/data/CleanBivData_1.27.20.RData")
#load(file="C:/Users/alpeterson7/Documents/MLH1repo/data/CleanBivData_1.27.20.RData")

@@

#this MLH1 only goes up to batch16!
#load(file="C:/Users/alpeterson7/Documents/MLH1repo/data/MLH1/MLH1_data_setup_12.20.19.RData")#batch18
#load(file="C:/Users/alpeterson7/Documents/MLH1repo/data/MLH1/MLH1_data_setup_12.20.19.RData")#fixing bugs, formating

@@

```{r male.show.plots, echo=FALSE, fig.width = 4, fig.height=2.5, warning=FALSE, include=FALSE}
#try customize the plot size

@@

#just.male_mouse.av <-  male_mouse.av %>% filter(sex== "male") %>% #filter(species == "M.musculus")

just.male_mouse.av <-  fomo.male_mouse.av %>% filter(sex== "male") %>% filter(species == "M.musculus")

just.male_mouse.av$sex <- factor(just.male_mouse.av$sex, ordered = FALSE)
just.male_mouse.av$strain <- factor(just.male_mouse.av$strain, ordered = FALSE)
just.male_mouse.av$subsp <- factor(just.male_mouse.av$subsp, ordered = FALSE)
just.male_mouse.av$species <- factor(just.male_mouse.av$species, ordered = FALSE)

@@

Musc.male_mouse.av <-  just.male_mouse.av %>% filter(sex== "male") %>% filter(subsp == "Musc")

@@

male.plot <- ggplot(just.male_mouse.av, aes(x = as.factor(strain), y = mean_co, color= strain))+geom_jitter( alpha=0.5, width = .2) +geom_boxplot()+ ggtitle("Male mouse mean")+scale_color_manual(values=colors_of_strains)+ylim(c(20,34))+theme(legend.position="none",
  panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
  panel.background = element_blank(), axis.line=element_line(colour = "black"))+theme(axis.text.x = element_text(angle = 45, hjust = 1))
male.plot

musc.male.plot <- ggplot(Musc.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Musc male of MLH1 distributions")+scale_color_manual(values=colors_of_strains)+ylim(c(20,34))

@@

@@

```{r Q2.totSC.glm, echo=FALSE}

Q2.glm_mus.av_totSC_M4 <- glm(mean.skel ~ subsp * strain, data= totSC.mouse.av.table %>% filter(sex == "male"))
sum.Q2.glm_mus.av_totSC_M4  <- summary(Q2.glm_mus.av_totSC_M1)$coefficients

#sum.Q2.glm_mus.av_totSC_M4$coefficients[,4]   pvalues of all effects
#sum.Q2.glm_mus.av_totSC_M4$coefficients[,1]  coefficients of all effects
#round(sum.Q2.glm_mus.av_totSC_M4$coefficients[,c(1,4)], 6)

#M4.2? (removed subspecies?)

Q2.glm_mus.av_totSC_M4.2 <- glm(mean.skel ~ strain, data= totSC.mouse.av.table %>% filter(sex == "male"))
sum.Q2.glm_mus.av_totSC_M4.2  <- summary(Q2.glm_mus.av_totSC_M4.2)
#round(summary(sum.Q2.glm_mus.av_totSC_M2)$coefficients[,c(1,4)], 6)

totSC.mouse.av.table <- add_strain(totSC.mouse.av.table)
totSC.mouse.av.table <- add_sex(totSC.mouse.av.table)
totSC.mouse.av.table <- add_subsp(totSC.mouse.av.table)
#within subspecies values

#Dom
Q2.glm_mus.av_totSC_M4.2_DOM <- glm(mean.skel ~ strain, data= totSC.mouse.av.table %>% filter(sex == "male") %>% filter(subsp == "Dom"))
sum.Q2.glm_mus.av_totSC_M4.2_DOM  <- summary(Q2.glm_mus.av_totSC_M4.2_DOM)
#there are significant strain effects in DOM,

#Musc
Q2.glm_mus.av_totSC_M4.2_MUSC <- glm(mean.skel ~ strain, data= totSC.mouse.av.table %>% filter(sex == "male") %>% filter(subsp == "Musc") )
sum.Q2.glm_mus.av_totSC_M4.2_MUSC  <- summary(Q2.glm_mus.av_totSC_M4.2_MUSC)
#KAZ -- is the only significant effect

#Mol
Q2.glm_mus.av_totSC_M4.2_MOL <- glm(mean.skel ~ strain, data= totSC.mouse.av.table %>% filter(sex == "male") %>% filter(subsp == "Mol") )
sum.Q2.glm_mus.av_totSC_M4.2_MOL  <- summary(Q2.glm_mus.av_totSC_M4.2_MOL)

@@

table( droplevels(fomo.male_mouse.av$strain) )
#there are cast strains too

Musc.male_mouse.av <-  fomo.male_mouse.av %>% filter(sex== "male") %>% filter(subsp == "Musc")

musc.male.plot <- ggplot(Musc.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Musc male of MLH1 distributions")+scale_color_manual(values=colors_of_strains)+ylim(c(20,34))
musc.male.plot

@@

Confirming the basic predictions, there is a positive correlation between mean MLH1 foci per cell and total SC ( Spearmans' r = `r cor( totSC.mouse.av.table$mean.skel, totSC.mouse.av.table$mean.MLH1, method = c("spearman") )`; p = `r cor.test(totSC.mouse.av.table$mean.skel, totSC.mouse.av.table$mean.MLH1, method=c("spearman") )$p.value`). Nevertheless, mean total SC only partially differentiates high-recombination and low-recombination strains (Figure X).  **Bret's note on positive correlation, 'A correlation across mice, across strains, or both?** While high-recombination strains have significantly more SC area in the total dataset (t-test; p = 0.01 -- `r Q2tot.sc_t.test_High.Low$p.value`), separate tests by strain show that only within _molossinus_ are high- and low-recombination strains significantly different (t-test; molossinus: p = 0.03 -- `r Q2tot.sc_t.test_High.Low_Mol$p.value`; musculus: p = 0.87  -- `r Q2tot.sc_t.test_High.Low_Musc$p.value`). 
**the values above are diferent between editions**

@@

**the values between versions are slightly different**
In a general linear model with total SC as the dependent variable, two subspecies effects are significant (p = musculus 1.2410^{-6} -- `r sum.Q2.glm_mus.av_totSC_M1[,4][2]`, molossinus p = 10^{-6} -- `r sum.Q2.glm_mus.av_totSC_M1[,4][3]`). In general linear models with reduced bivalent means as dependent variables, several subspecies and strain effects reach significance (p < 0.05) but they are not consistent across models, indicating to some extent the chromatin compaction evolution is decoupled from evolution in mean MLH1 foci per cell.   **Bret's comment; How can you differentiate between this biological conclusion and uncertainty due to the size of the datasets and statistical approaches?**

@@

>>>>

#load(file = here("data", "MLH1_data_setup_4.21.20.RData") )

load(file = "~./MLH1repo/data/Results_setup_4.21.20.RData")

@@

@@

```{r male.show.plots, echo=FALSE, fig.width = 4, fig.height=2.5, warning=FALSE}
#try customize the plot size

@@

#just.male_mouse.av <-  male_mouse.av %>% filter(sex== "male") %>% #filter(species == "M.musculus") 
#just.male_mouse.av <-  fomo.male_mouse.av %>% filter(sex== "male") %>% filter(species == "M.musculus")

#just.male_mouse.av$sex <- factor(just.male_mouse.av$sex, ordered = FALSE)
#just.male_mouse.av$strain <- factor(just.male_mouse.av$strain, ordered = FALSE)
#just.male_mouse.av$subsp <- factor(just.male_mouse.av$subsp, ordered = FALSE)
#just.male_mouse.av$species <- factor(just.male_mouse.av$species, ordered = FALSE)

@@

#Musc.male_mouse.av <-  just.male_mouse.av %>% filter(sex== "male") %>% filter(subsp == "Musc")

@@

#male.plot <- ggplot(just.male_mouse.av, aes(x = as.factor(strain), y = mean_co, color= strain))+geom_jitter( alpha=0.5, width = .2) +geom_boxplot()+ ggtitle("Male mouse mean")+scale_color_manual(values=colors_of_strains)+ylim(c(20,34))+theme(legend.position="none",
#  panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
#  panel.background = element_blank(), axis.line=element_line(colour = "black"))+theme(axis.text.x = element_text(angle = 45, hjust = 1))
#male.plot

#musc.male.plot <- ggplot(Musc.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Musc male of MLH1 distributions")+scale_color_manual(values=colors_of_strains)+ylim(c(20,34))

@@

```{r Q1.IFD.triangle, echo=FALSE}

#make triangle for the IFD measures (try doing female top male bottom)
#Curated_BivData[Curated_BivData$hand.foci.count == 2,]

#remove grey background
#add diagonal
#

IFD.Male.DOM.top.tri.plot <- ggplot( (Curated_BivData %>% filter(sex == "male") %>% filter(hand.foci.count == 2) %>% filter(subsp == "Dom") ),
  aes(y = rev(PER_Foci_2), x= PER_Foci_1, color=strain))+
  geom_jitter(alpha=.4)+
  ggtitle("Normalized IFD\nDom")+
  scale_color_manual(values=colors_of_strains)+xlim(c(0,1))+ylim(c(0,1))+
  theme(legend.position="none")+theme_classic()+geom_abline(intercept = 0, slope = 1)

IFD.Male.DOM.tri.plot

basey.base <- plot( y = (Curated_BivData %>% filter(sex == "male") %>% filter(hand.foci.count == 2) %>% filter(subsp == "Dom") )$PER_Foci_2, x= (Curated_BivData %>% filter(sex == "male") %>% filter(hand.foci.count == 2) %>% filter(subsp == "Dom") )$PER_Foci_1 )

```{r Q1.show.IFD_female.male.comparison, echo=FALSE}

#compare female IFD patterns which have telomeric bias (more similar to males)

@@

@@

#table( droplevels(fomo.male_mouse.av$strain) )
#there are cast strains too

#Musc.male_mouse.av <-  fomo.male_mouse.av %>% filter(sex== "male") %>% filter(subsp == "Musc")

#musc.male.plot <- ggplot( ( male_mouse.av %>% filter(species == "M.musculus") ), aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = mean_co, color= strain, alpha=0.5), width = .2 )+ ggtitle("Musc male of MLH1 distributions")+scale_color_manual(values=colors_of_strains)+ylim(c(20,34))
#musc.male.plot

@@

Confirming the basic predictions, there is a positive correlation between mean MLH1 foci per cell and total SC ( Spearmans' r = `r cor( totSC.mouse.av.table$mean.skel, totSC.mouse.av.table$mean.MLH1, method = c("spearman") )`; 
<no mean.MLH1>

p = `r cor.test(totSC.mouse.av.table$mean.skel, totSC.mouse.av.table$mean.MLH1, method=c("spearman") )$p.value`). Nevertheless, mean total SC only partially differentiates high-recombination and low-recombination strains (Figure X).  **Bret's note on positive correlation, 'A correlation across mice, across strains, or both?** While high-recombination strains have significantly more SC area in the total dataset (t-test; p = 0.01 -- `r Q2tot.sc_t.test_High.Low$p.value`), separate tests by strain show that only within _molossinus_ are high- and low-recombination strains significantly different (t-test; molossinus: p = 0.03 -- `r Q2tot.sc_t.test_High.Low_Mol$p.value`; musculus: p = 0.87  -- `r Q2tot.sc_t.test_High.Low_Musc$p.value`). 
**the values above are diferent between editions**

@@

**the values between versions are slightly different**
In a general linear model with total SC as the dependent variable, two subspecies effects are significant (p = musculus 1.2410^{-6} -- `r sum.Q2.glm_mus.av_totSC_M4[,4][2]`, molossinus p = 10^{-6} -- `r sum.Q2.glm_mus.av_totSC_M4[,4][3]`). In general linear models with reduced bivalent means as dependent variables, several subspecies and strain effects reach significance (p < 0.05) but they are not consistent across models, indicating to some extent the chromatin compaction evolution is decoupled from evolution in mean MLH1 foci per cell.   **Bret's comment; How can you differentiate between this biological conclusion and uncertainty due to the size of the datasets and statistical approaches?**

@@

```{r Q2.triangle.IFD, echo=FALSE}

### 2020-04-22 src/Result_Setup.Rmd
fixing bugs and playing with triangle plots

```{r MLH1.HetC.rates, echo=FALSE, include=FALSE, warning=FALSE}
#calculate the initial HetC levels from best quality mice and cells (compare both levels (cell and mouse average levels))
#subset MLH1_DF, remove non passing mice (as diagnoised from plots)
#remove quality 5 cells

@@

```{r tot.skel.table, echo=FALSE}
totSC.mouse.av.table <- ddply(.data=MLH1.merge.Skel, 
                 .(mouse),
                 summarize, 
                 ncells = length(unique(Original.Name)),
                 nmice = length(unique(mouse)),
                # mean.MLH1 = mean( nMLH1.foci, na.rm = TRUE),
                # var.MLH1 = var( nMLH1.foci, na.rm = TRUE),

@@

totSC.strain.av.table <- ddply(.data=totSC.mouse.av.table, 
                 .(sex,strain),
                 summarize, 
                 ncells = sum(ncells),
                 nmice = length(unique(mouse)),
                # mean.MLH1 = mean( nMLH1.foci, na.rm = TRUE),
                # var.MLH1 = var( nMLH1.foci, na.rm = TRUE),

@@

Q2.glm_mus.av_totSC_M4 <- glm(mean.skel ~ subsp * strain, data= totSC.mouse.av.table %>% filter(sex == "male"))
sum.Q2.glm_mus.av_totSC_M4  <- summary(Q2.glm_mus.av_totSC_M1)$coefficients

@@

Q2.glm_mus.av_1COpos_M4 <- glm(norm.first.foci.pos ~ subsp * strain, data= mouse.avs_4MM %>% filter(sex == "male"))
sum.Q2.glm_mus.av_1COpos_M1  <- summary(Q2.glm_mus.av_1COpos_M1)$coefficients
#summary(Q2.glm_mus.av_1COpos_M1)$coefficients[,4]

#with.chrm.legth  ?
Q2.glm_mus.av_1COpos_M4 <- glm(norm.first.foci.pos ~ subsp * strain, data= mouse.avs_4MM %>% filter(sex == "male"))

@@

```{r Q2.1COpos.show.MOL, echo=FALSE}

#(mouse.avs_4MM %>%  filter(sex == "male") %>%  filter(subsp == "Mol") %>% filter(Rec.group == #"0"))$norm.first.foci.pos

mus.av1COpos.plot <- ggplot(data= mouse.avs_4MM %>%  filter(sex == "male") %>%  filter(subsp == "Mol"),
                      aes(x= mouse, y = norm.first.foci.pos))+geom_point()

#+facet_wrap(~mouse, scales = "free")

single.biv_1COpos.plot_Mol <- ggplot(data= Curated_BivData %>%  filter(sex == "male") %>%  filter(subsp == "Mol") %>%  filter(hand.foci.count == "1") %>%  filter(chromosomeLength < 100),
                      aes(x= fileName, y = PER_Foci_1, color=chromosomeLength))+geom_point()+facet_wrap(~mouse, scales = "free") + ylim(c(0,1))

#+scale_color_manual(values=colors_of_strains)

single.biv_1COpos.plot_Dom <- ggplot(data= Curated_BivData %>% filter(sex == "male") %>%filter(subsp == "Dom") %>%  filter(hand.foci.count == "1"),
                      aes(x= fileName, y = PER_Foci_1, color=strain ))+geom_point()+facet_wrap(~mouse, scales = "free") + ylim(c(0,1))+scale_color_manual(values=colors_of_strains)

#CHRM.length is very significant-- some of the same subsp and strains are sig     
```

@@

#M4.ABS
Q2.M4glm_IFD.raw <- glm(mean_IFD.2CO_ABS ~ subsp * strain, data= mouse.avs_4MM %>% filter(sex == "male")    )
sum.Q2.M1glm_IFD.raw <- summary(Q2.M4glm_IFD.raw)

@@

#M4.2.raw
Q2.M4.2glm_IFD.raw <- glm(mean_IFD.2CO_ABS ~ strain, data= mouse.avs_4MM %>% filter(sex == "male"))
Sum.Q2.M2glm_IFD.raw <- summary(Q2.M2glm_IFD.raw)

@@

#within subspecies
#Dom
Q2.glm.M2_IFD_Dom <- glm(mean_IFD.2CO_ABS ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Dom",])
#summary(Q2.glm.M2_IFD_Dom)

#Musc
Q2.glm.M2_IFD_Musc <- glm(mean_IFD.2CO_ABS ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",])
#summary(Q2.glm.M2_IFD_Musc)

#Mol
Q2.glm.M2_IFD_Mol <- glm(mean_IFD.2CO_ABS ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",])
#summary(Q2.glm.M2_IFD_Mol)

#Dom
Q2.glm.M2_IFD.PER_Dom <- glm(mean_IFD.2CO_PER ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Dom",])
#summary(Q2.glm.M2_IFD.PER_Dom)

#Musc
Q2.glm.M2_IFD.PER_Musc <- glm(mean_IFD.2CO_PER ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",])
#summary(Q2.glm.M2_IFD.PER_Musc)

#Mol
Q2.glm.M2_IFD.PER_Mol <- glm(mean_IFD.2CO_PER ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",])
#summary(Q2.glm.M2_IFD.PER_Mol)

@@

```{r Save.File, echo=FALSE}
#save Rdata
save.image("~./MLH1repo/data/Results_setup_4.21.20.RData")
#save.image("data/BivData_4_PCA.RData")
```

>>>>

```{r MLH1.HetC.rates, echo=FALSE, warning=FALSE}
#calculate the initial HetC levels from best quality mice and cells (compare both levels (cell and mouse average levels))
#subset MLH1_DF, remove non passing mice (as diagnoised from plots)
#remove quality 5 cells

@@

```{r tot.skel.table, echo=FALSE}

MLH1.merge.Skel$nMLH1.foci <- as.numeric(MLH1.merge.Skel$nMLH1.foci)

totSC.mouse.av.table <- ddply(.data=MLH1.merge.Skel, 
                 .(mouse),
                 summarize, 
                 ncells = length(unique(Original.Name)),
                 nmice = length(unique(mouse)),

                 mean.MLH1 = mean(nMLH1.foci, na.rm = TRUE),
                 var.MLH1 = var( nMLH1.foci, na.rm = TRUE),

@@

totSC.strain.av.table <- ddply(.data=totSC.mouse.av.table, 
                 .(sex,strain),
                 summarize, 
                 ncells = sum(ncells),
                 nmice = length(unique(mouse)),

                av.MLH1 = mean( mean.MLH1, na.rm = TRUE),
                # var.MLH1 = var( nMLH1.foci, na.rm = TRUE),

@@

Q2.glm_mus.av_totSC_M4 <- glm(mean.skel ~ subsp * strain, data= totSC.mouse.av.table %>% filter(sex == "male"))
sum.Q2.glm_mus.av_totSC_M4  <- summary(Q2.glm_mus.av_totSC_M4)$coefficients

@@

Q2.glm_mus.av_1COpos_M4 <- glm(norm.first.foci.pos ~ subsp * strain, data= mouse.avs_4MM %>% filter(sex == "male"))
sum.Q2.glm_mus.av_1COpos_M4  <- summary(Q2.glm_mus.av_1COpos_M4)$coefficients
#summary(Q2.glm_mus.av_1COpos_M1)$coefficients[,4]

#with.chrm.legth  ?
#Q2.glm_mus.av_1COpos_M4 <- glm(norm.first.foci.pos ~ subsp * strain, data= mouse.avs_4MM %>% filter(sex == "male"))

@@

@@

#M4.ABS
Q2.M4glm_IFD.raw <- glm(mean_IFD.2CO_ABS ~ subsp * strain, data= mouse.avs_4MM %>% filter(sex == "male")    )
sum.Q2.M4glm_IFD.raw <- summary(Q2.M4glm_IFD.raw)

@@

#M4.2.raw
Q2.M4.2glm_IFD.raw <- glm(mean_IFD.2CO_ABS ~ strain, data= mouse.avs_4MM %>% filter(sex == "male"))
Sum.Q2.M4.2glm_IFD.raw <- summary(Q2.M4.2glm_IFD.raw)

@@

#within subspecies
#Dom
Q2.glm.M4.2_IFD_Dom <- glm(mean_IFD.2CO_ABS ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Dom",])
#summary(Q2.glm.M4.2_IFD_Dom)

#Musc
Q2.glm.M4.2_IFD_Musc <- glm(mean_IFD.2CO_ABS ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",])
#summary(Q2.glm.M4.2_IFD_Musc)

#Mol
Q2.glm.M4.2_IFD_Mol <- glm(mean_IFD.2CO_ABS ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",])
#summary(Q2.glm.M4.2_IFD_Mol)

#Dom
Q2.glm.M4.2_IFD.PER_Dom <- glm(mean_IFD.2CO_PER ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Dom",])
#summary(Q2.glm.M4.2_IFD.PER_Dom)

#Musc
Q2.glm.M4.2_IFD.PER_Musc <- glm(mean_IFD.2CO_PER ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",])
#summary(Q2.glm.M4.2_IFD.PER_Musc)

#Mol
Q2.glm.M4.2_IFD.PER_Mol <- glm(mean_IFD.2CO_PER ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",])
#summary(Q2.glm.M4.2_IFD.PER_Mol)
```

@@

```{r Save.File, echo=FALSE}
#save Rdata
save.image("~./MLH1repo/data/Results_setup_4.21.20.RData")
